---
title: "Bayesian Estimation with brms"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bayesian Estimation with brms}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  eval = FALSE
)
```

## Overview

This vignette demonstrates a complete Bayesian estimation workflow using `vrm` and `brms`, modeled after the post-estimation analysis in Weber (2025). The workflow covers model fitting across multiple families (linear, ordinal, multinomial), posterior prediction, marginal effects, counterfactual inference via g-computation, and publication-quality visualization. All examples use the 2020 Western States Survey (`wss20`), a natural experiment in which respondents were surveyed before and after the presidential election.

## Data and Research Design

The Western States Survey is a pre-post election design. The key treatment variable (`prepost`) indicates whether a respondent completed the survey before or after the 2020 election. The moderator (`vote_trump`) captures whether the respondent voted for Trump. The outcome is support for various electoral contestation behaviors (e.g., attending a march, burning a flag, requesting a recount), each measured on a 5-point Likert scale.

```{r load-data}
library(vrm)
library(brms)
library(tidybayes)
library(dplyr)
library(tidyr)

data(wss20, package = "vrm")
```

## Step 1: IPTW Weights (Optional)

Before fitting outcome models, we can estimate inverse probability of treatment weights (IPTW) to adjust for pre-treatment covariate imbalance. `ps_weight()` fits a propensity score model for the treatment assignment and returns stabilized weights. It supports logistic regression (`"glm"`), gradient boosted models (`"gbm"`), and random forest (`"rf"`).

```{r iptw-weights}
ps_result <- ps_weight(
  data      = wss20,
  treatment = "prepost",
  formula   = ~ female + latino + college + scale(age),
  methods   = "glm"
)

# ps_weight() returns a list: $data has new ps/wts columns, $fits has the model
wss20 <- ps_result$data

# Check weight distribution
summary(wss20$wts)
```

When multiple methods are requested, columns are suffixed (e.g., `wts_glm`, `wts_gbm`, `wts_rf`):

```{r iptw-multi}
ps_multi <- ps_weight(
  data      = wss20,
  treatment = "prepost",
  formula   = ~ female + latino + college + scale(age),
  methods   = c("glm", "gbm")
)
wss20 <- ps_multi$data
# Now has: ps_glm, wts_glm, ps_gbm, wts_gbm
```

## Step 2: Model Fitting

`vrm` just consists of a lot of wrappers. There's not a lot original
here, but just useful things that I
keep cut-and-pasting across projects.

So, the idea is to build a set of functions from different project
analysis scripts I've run over time,
and just put them into one place, this package.

`brms::brm()` that accept formula components as character strings. This makes it easy to fit the same specification across different families.

### Linear Model

A Gaussian (linear) model...It doesn't make sense in all circumstances, but usually a fine place to start.

```{r fit-linear}
fit_linear <- brms.linear(
  data      = wss20,
  DV        = "contestation_value",
  treatment = "prepost",
  moderator = "vote_trump",
  controls  = "female + latino + college + scale(age)",
  cores = 4, chains = 4, iter = 2000
)
```

### Weighted Linear Model (IPTW)

To incorporate IPTW weights, pass the weight column name via `weights`. The function builds `DV | weights(wts) ~ ...` internally:

```{r fit-linear-iptw}
fit_linear_w <- brms.linear(
  data      = wss20,
  DV        = "contestation_value",
  treatment = "prepost",
  moderator = "vote_trump",
  controls  = "female + latino + college + scale(age)",
  cores = 4, chains = 4, iter = 2000
)
```

### Ordinal Model

A cumulative logit model assumes an ordinal response, and cumulative probabilities. The covariate(s) effects are proportional across response categories (i.e., parallel lines assumption).

```{r fit-ordinal}
fit_ord <- brms.ordinal(
  data      = wss20,
  DV        = "contestation_value",
  treatment = "prepost",
  moderator = "vote_trump",
  controls  = "female + latino + college + scale(age)",
  cores = 4, chains = 4, iter = 2000
)
```

### Multinomial Model

A categorical (multinomial) model relaxes this...response categories are nominal, and covariate effects are estimated separately for each category (relative to a reference category). 

```{r fit-nominal}
fit_mlogit <- brms.nominal(
  data      = wss20,
  DV        = "contestation_value",
  treatment = "prepost",
  moderator = "vote_trump",
  controls  = "female + latino + college + scale(age)",
  cores = 4, chains = 4, iter = 2000
)
```

### General-Purpose Wrapper

Alternatively, `brms_fit()` accepts any `brms` family:

```{r brms-fit}
fit_alt <- brms_fit(
  data      = wss20,
  DV        = "contestation_value",
  treatment = "prepost",
  moderator = "vote_trump",
  controls  = "female + latino + college + scale(age)",
  family    = cumulative("logit")
)
```

### Fitting a Model Battery

Fitting the same model across model items -- one model at a time -- becomes tedious. Use `fit_brms_battery()` to fit a battery of models across multiple items and families.

The `wss20` data-builder creates `tmp_contestation_*` columns in wide format before reshaping to long. To use the battery, work with the wide-format data (i.e., before `pivot_longer`):

```{r battery}
items <- c(
  "tmp_contestation_attend_march",
  "tmp_contestation_criticize_election",
  "tmp_contestation_burn_flag",
  "tmp_contestation_recount",
  "tmp_contestation_court"
)
families <- list(ord = cumulative(), linear = gaussian())

# Unweighted battery (use wide-format wss20 before pivot_longer)
models <- fit_brms_battery(
  data     = wss20_wide,
  items    = items,
  rhs      = "prepost * vote_trump + female + latino + college + scale(age)",
  families = families,
  data_dir = tempdir(),
  prefix   = "wss20",
  weight_var = NULL,
  cores = 4, chains = 3, iter = 2000
)
```

To run the battery with IPTW weights, supply `weight_var`. This fits both unweighted and weighted variants for each item x family combination, using `DV | weights(wts) ~ rhs` in `brms`:

```{r battery-iptw}
models_w <- fit_brms_battery(
  data       = wss20_wide,
  items      = items,
  rhs        = "prepost * vote_trump + female + latino + college + scale(age)",
  families   = families,
  data_dir   = tempdir(),
  prefix     = "wss20",
  weight_var = "wts",
  cores = 4, chains = 3, iter = 2000
)

# Model names include a _w suffix for weighted variants
names(models_w)
```

Alternatively, with the long-format `wss20`, fit a single model on `contestation_value` and include `contestation` as a covariate:

```{r battery-long}
fit_long <- brms.linear(
  data      = wss20,
  DV        = "contestation_value",
  treatment = "prepost",
  moderator = "vote_trump",
  controls  = "female + latino + college + scale(age) + contestation",
  cores = 4, chains = 4, iter = 2000
)
```

## Step 3: Model Diagnostics

Check convergence across the battery:

```{r diagnostics}
diag_tbl <- summarize_brm_diagnostics(models)
print_diagnostic_warnings(diag_tbl, label = "WSS 2020")

# Inspect the diagnostic table
diag_tbl
```

This reports divergent transitions, max treedepth exceedances, Rhat, and effective sample size ratios for each model, flagging any with potential issues.

## Step 4: Posterior Predicted Means

`posterior_means()` computes posterior predicted means across a grid of treatment and moderator values, holding all other covariates at their observed means.

```{r posterior-means}
pred_means <- posterior_means(
  model = fit_linear,
  xvar  = "prepost",
  mvar  = "vote_trump",
  xval  = c(0, 1),
  mval  = c(0, 1)
)
pred_means
```

For ordinal models, the output includes a `.category` column with predicted probabilities for each response level:

```{r posterior-means-ord}
pred_ord <- posterior_means(
  model = fit_ord,
  xvar  = "prepost",
  mvar  = "vote_trump",
  xval  = c(0, 1),
  mval  = c(0, 1)
)
pred_ord
```

## Step 5: Posterior Marginal Effects

`posterior_pme()` computes the average marginal effect of the treatment variable, conditioned on the moderator. This is the difference in predicted values between `xrange[2]` and `xrange[1]`, evaluated at each level of the moderator.

```{r marginal-effects}
me_linear <- posterior_pme(
  model  = fit_linear,
  xvar   = "prepost",
  mvar   = "vote_trump",
  xrange = c(0, 1),
  mrange = c(0, 1)
)
me_linear
```

For ordinal models, marginal effects are computed per response category:

```{r marginal-effects-ord}
me_ord <- posterior_pme(
  model  = fit_ord,
  xvar   = "prepost",
  mvar   = "vote_trump",
  xrange = c(0, 1),
  mrange = c(0, 1)
)
me_ord
```

## Step 6: Posterior Predictive Differences

`posterior_pred_diff()` provides a cleaner interface for computing the posterior distribution of contrasts (e.g., post minus pre), with full control over covariate profiles:

```{r pred-diff}
diff_linear <- posterior_pred_diff(
  model = fit_linear,
  xvar  = "prepost",
  mvar  = "vote_trump",
  xval  = c(0, 1),
  mval  = c(0, 1),
  prob  = 0.95
)
diff_linear

# For ordinal models, filter to a specific category
diff_ord_support <- posterior_pred_diff(
  model    = fit_ord,
  xvar     = "prepost",
  mvar     = "vote_trump",
  xval     = c(0, 1),
  mval     = c(0, 1),
  category = "5"
)
diff_ord_support
```

## Step 7: Raw Posterior Draws

For custom visualizations (ridgeplots, density plots), extract the full posterior distribution with `posterior_draws_grid()`:

```{r draws}
draws <- posterior_draws_grid(
  model  = fit_linear,
  xvar   = "prepost",
  mvar   = "vote_trump",
  xval   = c(0, 1),
  mval   = c(0, 1),
  ndraws = 500
)
head(draws)
```

You can also set specific covariate profiles:

```{r draws-covariates}
draws_profile <- posterior_draws_grid(
  model      = fit_linear,
  xvar       = "prepost",
  mvar       = "vote_trump",
  xval       = c(0, 1),
  mval       = c(0, 1),
  covariates = list(female = 1, college = 1),
  ndraws     = 500
)
```

## Step 8: Counterfactual Inference (G-Computation)

G-computation estimates individual-level treatment effects by predicting outcomes under both observed and counterfactual treatment conditions. This three-step workflow is the core of the causal inference framework.

### Step 8a: Generate Counterfactual Data

`generate_counterfactual_data()` doubles the dataset, creating one copy with actual treatment values and one with flipped values:

```{r counterfactual}
cf_data <- generate_counterfactual_data(
  data          = wss20,
  treatment_var = "prepost"
)
# Each respondent now appears twice
table(cf_data$pre_condition)
```

### Step 8b: Obtain Posterior Predictions

Feed the counterfactual data through `tidybayes::add_epred_draws()` to get posterior predictions under both conditions:

```{r cf-predictions}
cf_preds <- cf_data %>%
  tidybayes::add_epred_draws(fit_linear, ndraws = 500)
```

### Step 8c: Calculate Individual Treatment Effects

`calculate_individual_effects()` pivots from long to wide on condition, computes the per-individual difference, and applies direction correction:

```{r individual-effects}
ind_effects <- calculate_individual_effects(
  predictions_long = cf_preds,
  pred_col         = ".epred",
  treatment_var    = "prepost",
  group_vars       = c("vote_trump")
)
head(ind_effects)
```

### Step 8d: Summarize Treatment Effects

Aggregate to conditional average treatment effects (CATEs):

```{r summarize-effects}
cate <- summarize_treatment_effects(
  effects_data = ind_effects,
  effect_var   = "treatment_effect",
  group_vars   = c("vote_trump"),
  conf_level   = 0.95,
  draw_level   = TRUE
)
cate
```

For ordinal models, include `.category` in the grouping:

```{r summarize-effects-ord}
cf_preds_ord <- cf_data %>%
  tidybayes::add_epred_draws(fit_ord, ndraws = 500)

ind_effects_ord <- calculate_individual_effects(
  predictions_long = cf_preds_ord,
  pred_col         = ".epred",
  treatment_var    = "prepost",
  group_vars       = c("vote_trump", ".category")
)

cate_ord <- summarize_treatment_effects(
  effects_data = ind_effects_ord,
  effect_var   = "treatment_effect",
  group_vars   = c("vote_trump", ".category"),
  draw_level   = TRUE
)
cate_ord
```

## Step 9: Marginal Effects Battery

When you have fitted a battery of models, `marginal_effects_battery()` computes marginal effects across all items and families at once:

```{r me-battery}
me_all <- marginal_effects_battery(
  model_list = models,
  items      = items,
  treatment  = "prepost",
  moderator  = "vote_trump"
)
me_all
```

## Step 10: Publication Tables

`brms_regression_table()` produces a formatted `gt` table from a model list:

```{r reg-table}
level_lookup <- list(
  attend_march       = c("Strongly Oppose", "Oppose", "Neutral", "Support", "Strongly Support"),
  burn_flag          = c("Strongly Oppose", "Oppose", "Neutral", "Support", "Strongly Support"),
  recount            = c("Strongly Oppose", "Oppose", "Neutral", "Support", "Strongly Support"),
  court              = c("Strongly Oppose", "Oppose", "Neutral", "Support", "Strongly Support"),
  criticize_election = c("Strongly Oppose", "Oppose", "Neutral", "Support", "Strongly Support")
)

brms_regression_table(
  model_list    = models,
  dataset_label = "2020 Western States Survey",
  level_lookup  = level_lookup
)
```

## Step 11: Visualization

### Ridgeplots

Combine draws across items and visualize with `gg_posterior_ridges()`:

```{r ridgeplot}
library(purrr)
library(ggridges)

all_draws <- map_dfr(items, function(it) {
  posterior_draws_grid(
    model  = models[[paste0(it, "_linear")]],
    xvar   = "prepost",
    mvar   = "vote_trump",
    xval   = c(0, 1),
    mval   = c(0, 1),
    ndraws = 500
  ) %>%
    mutate(item = it)
})

gg_posterior_ridges(
  draws        = all_draws,
  ridge_var    = "item",
  fill_var     = "prepost",
  facet_var    = "vote_trump",
  fill_labels  = c("0" = "Pre-Election", "1" = "Post-Election"),
  facet_labels = c("0" = "Biden Voter", "1" = "Trump Voter"),
  title        = "Posterior Predicted Mean Support (Linear)"
)
```

### Forest Plot of Predictive Differences

```{r forest-plot}
all_diffs <- map_dfr(items, function(it) {
  posterior_pred_diff(
    model = models[[paste0(it, "_linear")]],
    xvar  = "prepost",
    mvar  = "vote_trump"
  ) %>%
    mutate(item = it)
})

gg_pred_diff(
  data         = all_diffs,
  x_var        = "item",
  color_var    = "vote_trump",
  color_labels = c("0" = "Biden Voter", "1" = "Trump Voter"),
  title        = "Posterior Predictive Difference (Post − Pre)"
)
```

### Marginal Effects (Ordinal Categories)

```{r marginal-plot}
ggMargins_categories(
  data               = me_ord,
  mvar               = "vote_trump",
  title              = "Marginal Effect of Post-Election Period",
  point_color_values = c("#E63946", "#457B9D")
)
```

### Battery Forest Plot

```{r battery-forest}
plot_marginal_effects(
  me_data = me_all,
  title   = "Marginal Effect of Post-Election Period",
  colors  = c("Trump" = "#E63946", "Non-Trump" = "#457B9D")
)
```

## Summary

The `vrm` Bayesian workflow:

1. **Weight** (optional) with `ps_weight()` for IPTW-adjusted estimation
2. **Fit** models with `brms.linear()`, `brms.ordinal()`, `brms.nominal()`, or `fit_brms_battery()`
3. **Diagnose** with `summarize_brm_diagnostics()`
4. **Predict** with `posterior_means()`, `posterior_pme()`, `posterior_pred_diff()`
5. **Extract draws** with `posterior_draws_grid()` for custom visualizations
6. **Counterfactual** inference with `generate_counterfactual_data()` + `calculate_individual_effects()` + `summarize_treatment_effects()`
7. **Tabulate** with `brms_regression_table()`
8. **Visualize** with `gg_posterior_ridges()`, `gg_pred_diff()`, `ggMargins_categories()`, `plot_marginal_effects()`
