---
title: "Building Survey Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Building Survey Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

```{r, echo = TRUE, cache = TRUE}
library(tidyverse)
library(readstata13)
library(haven)
library(broom)
library(gt)
pkg_root <- path.expand("~/Dropbox/github_repos/electoralContestation")
data_raw <- path.expand("~/Dropbox/electoral_contestation-main/Data Repository")
data_out <- file.path(pkg_root, "data")

# ── Load raw data ──
wss20 <- read.dta13(file.path(data_raw, "2020 WSS.dta")) |>
  mutate(
    date = as.Date(starttime, format = "%Y-%m-%d"),
    staticDate = mean(date, na.rm = TRUE)
  )

cesASU <- read.dta13(file.path(data_raw, "2022 CES ASU.dta")) |>
  mutate(
    date = as.Date(starttime, format = "%Y-%m-%d"),
    staticDate = mean(date, na.rm = TRUE)
  )

cesBYU24 <- read_sav(file.path(data_raw, "CCES24_BYU_OUTPUT.sav")) |>
  mutate(
    date = as.Date(starttime, format = "%Y-%m-%d"),
    staticDate = mean(date, na.rm = TRUE)
  )

wss24 <- read.dta13(file.path(data_raw, "2024 WSS.dta")) |>
  mutate(
    date = as.Date(starttime, format = "%Y-%m-%d"),
    staticDate = mean(date, na.rm = TRUE)
  )

cat(
  "Data files loaded:",
  "\n- wss20: ", nrow(wss20), " rows",
  "\n- cesASU: ", nrow(cesASU), " rows",
  "\n- cesBYU24: ", nrow(cesBYU24), " rows",
  "\n- wss24: ", nrow(wss24), " rows"
)
```

This vignette demonstrates how to use `vrm` recode utilities to build
analysis-ready survey data from raw survey files. The example uses Western
States Survey (WSS) data from 2020.

## 2020

### Sampling Frame

This sample includes respondents across five western states, with a Latino
oversample.

YouGov interviewed 3600 respondents who were then matched down to a sample of
3000 to produce the final dataset. An oversample of 600 Latinos were interviewed
for a total of 3600. The respondents were matched to a sampling frame on gender,
age, race, and education. The frame was constructed by stratified sampling from
the full 2018 American Community Survey (ACS) 1-year sample with selection
within strata by weighted sampling with replacements (using the person weights
on the public use file).

The matched cases were weighted to the sampling frame using propensity scores.
The matched cases and the frame were combined and a logistic regression was
estimated for inclusion in the frame. The propensity score function included
age, gender, race/ethnicity, years of education, and region. The propensity
scores were grouped into deciles of the estimated propensity score in the frame
and post-stratified according to these deciles.

The weights were then post-stratified on 2016 Presidential vote choice, and a
four-way stratification of gender, age (4-categories), race (4-categories), and
education (4-categories), to produce the final weight.

### Design

The empirical design is a natural experiment; comparable numbers of participants
complete pre or post election, or interregnum period. Five "core" items are
present:

- Attend march or demonstration
- Burn the American flag
- Contest the outcome in the courts
- Support ballot recounts
- Publicly criticize the integrity or fairness of the election

Two questions are experimentally varied:

- Attend march or demonstration ["even if turns violent"]
- Publicly criticize the integrity or fairness of the election ["on social media"]

### Computed Demographics

Variables that require transformations beyond simple value mapping stay in
`mutate()` — date arithmetic, `zero_one()` rescaling, and conditional logic.

```{r}
library(vrm)
library(dplyr)
library(tidyr)
library(stringr)

original_columns <- names(wss20)

wss20 <- wss20 %>%
  mutate(
    age = zero_one(2020 - birthyr),
    survey_weight = weight,
    caseidID = caseid,
    state = as.character(inputstate),
    year = 2020,
    survey = "wss20",
    DATE = as.Date(endtime, "%Y-%m-%d"),
    post_election = ifelse(DATE >= "2020-11-03", 1, 0),
    post_call = ifelse(DATE >= "2020-11-07", 1, 0),
    uncertainty = ifelse(DATE > "2020-11-03" & DATE < "2020-11-07", 1, 0),
    prepost = ifelse(DATE > "2020-11-03", "post", "pre"),
    # Party feeling thermometers (0–100 → 0–1)
    therm_republican = zero_one(as.numeric(WSS29_1)),
    therm_democrat = zero_one(as.numeric(WSS29_2)),
    # Ideology placement (0–100 → 0–1)
    ideo_rep_state = zero_one(as.numeric(WSS28_1)),
    ideo_rep_national = zero_one(as.numeric(WSS28_2)),
    ideo_dem_state = zero_one(as.numeric(WSS28_3)),
    ideo_dem_national = zero_one(as.numeric(WSS28_4)),
    ideo_self = zero_one(as.numeric(WSS28_5)),
    # AZ candidate ideology placement (0–100 → 0–1)
    az_ideo_kelly = zero_one(as.numeric(AZ312_1)),
    az_ideo_mcsally = zero_one(as.numeric(AZ313_1))
  )
```

### Item Recodes

All simple value-mapping recodes use `recodeList()`. Common coding vectors
from `vrm` (`five_r`, `five_n`, `four_r`, `four_n`, `binary_10`, `binary_01`)
eliminate repetitive `case_when()` blocks. Each thematic block runs
`recodeList()` independently and reports Cronbach's $\alpha$ for multi-item
scales.

#### Experimental Treatments

```{r experimental-treatments}
wss20 <- recodeList(wss20, list(
  list(column = "WSS33_split", recode_rules = c("2" = "National", "1" = 1, "0" = "State"), new_column = "efficacy_treat"),
  list(column = "WSS28_split", recode_rules = c("2" = "Elected Officials", "1" = "People"), new_column = "ideology_1_treat"),
  list(column = "WSS29_split", recode_rules = c("2" = "People_State", "1" = "People"), new_column = "ideology_2_treat"),
  list(column = "WSS40_1_split", recode_rules = c("1" = "Violent March", "2" = "March"), new_column = "violent_march_treat"),
  list(column = "WSS40_5_split", recode_rules = c("1" = "Social Media Criticism", "2" = "Criticism"), new_column = "socialmedia_criticism_treat"),
  list(column = "WSS47_split", recode_rules = c("1" = "Federal", "2" = "State_or_Local"), new_column = "covid_treat"),
  list(column = "WSS54_split", recode_rules = c("1" = "pre", "2" = "post"), new_column = "institution_treat"),
  list(column = "AZ_split", recode_rules = c("1" = "partisan", "2" = "hispanic"), new_column = "ft_arizona_treat"),
  list(column = "AZ310_split", recode_rules = c("1" = "control", "2" = "information"), new_column = "senate_az_treat"),
  list(column = "WSS40_1_split", recode_rules = binary_10, new_column = "violent_treat"),
  list(column = "WSS40_5_split", recode_rules = binary_10, new_column = "socialmedia_treat")
))
```

#### Demographics

```{r demographics}
wss20 <- recodeList(wss20, list(
  list(column = "race", recode_rules = c("1" = 0, "2" = 1, "3" = 0, "4" = 0, "5" = 0), new_column = "black"),
  list(column = "race", recode_rules = c("1" = 1, "2" = 0, "3" = 0, "4" = 0, "5" = 0), new_column = "white"),
  list(column = "race", recode_rules = c("1" = 0, "2" = 0, "3" = 1, "4" = 0, "5" = 0), new_column = "latino"),
  list(column = "race", recode_rules = c("1" = 0, "2" = 0, "3" = 0, "4" = 1, "5" = 0), new_column = "asian"),
  list(column = "race", recode_rules = c("1" = 0, "2" = 0, "3" = 0, "4" = 0, "5" = 1), new_column = "american_indian"),
  list(column = "gender", recode_rules = c("2" = 1, "1" = 0), new_column = "female"),
  list(column = "marstat", recode_rules = c("1" = 1, "2" = 0, "3" = 0, "4" = 0, "5" = 0, "6" = 0), new_column = "married"),
  list(column = "educ", recode_rules = c("1" = 0, "2" = 0, "3" = 0, "4" = 0, "5" = 1, "6" = 1), new_column = "college")
))
```

#### Most Important Problem

```{r mip}
wss20 <- recodeList(wss20, list(
  list(column = "WSS15s", recode_rules = c(
    "1"  = "environment",
    "2"  = "Police",
    "3"  = "Jobs/economy",
    "4"  = "housing",
    "5"  = "Immigration",
    "6"  = "Trade",
    "8"  = "Foreign Policy",
    "9"  = "Healthcare",
    "10" = "COVID",
    "11" = "Other"
  ), new_column = "most_important_problem")
))
```

#### Party Identification, Ideology, and Religion

```{r party-ideology}
wss20 <- recodeList(wss20, list(
  list(column = "pid7", recode_rules = c(
    "1" = 1, "2" = 1, "3" = 1, "4" = 2,
    "5" = 3, "6" = 3, "7" = 3
  ), new_column = "party_identification3"),
  list(column = "pid7", recode_rules = c(
    "1" = 1, "2" = 2, "3" = 3, "4" = 4,
    "5" = 5, "6" = 6, "7" = 7
  ), new_column = "party_identification7"),
  list(
    column = "ideo5", recode_rules = c(
      "1" = 1, "2" = 2, "3" = 3, "4" = 4, "5" = 5
    ),
    new_column = "ideology"
  ),
  list(column = "religpew", recode_rules = c(
    "1" = 1, "2" = 1, "3" = 1, "4" = 1,
    "5" = 0, "6" = 0, "7" = 0, "8" = 0, "9" = 0,
    "10" = 0, "11" = 0, "12" = 0
  ), new_column = "christian"),
  list(
    column = "pew_churatd", recode_rules = c(
      "1" = 6, "2" = 5, "3" = 4, "4" = 3, "5" = 2, "6" = 1
    ),
    new_column = "religiosity_attendance"
  ),
  list(column = "newsint", recode_rules = four_r, new_column = "political_interest"),
  list(column = "WSS06", recode_rules = seven_n, new_column = "risk_aversion")
))
```

#### Authoritarianism (4 items, $\alpha$)

```{r authoritarianism}
wss20 <- recodeList(wss20, list(
  list(column = "WSS07_1a", recode_rules = binary_01, new_column = "auth_1"),
  list(column = "WSS07_1b", recode_rules = binary_10, new_column = "auth_2"),
  list(column = "WSS07_1c", recode_rules = binary_01, new_column = "auth_3"),
  list(column = "WSS07_1d", recode_rules = binary_01, new_column = "auth_4")
))

wss20$authoritarianism <- rowMeans(
  cbind(wss20$auth_1, wss20$auth_2, wss20$auth_3, wss20$auth_4),
  na.rm = TRUE
)
psych::alpha(wss20[, c("auth_1", "auth_2", "auth_3", "auth_4")])
```

#### Social Dominance Orientation (4 items, $\alpha$)

```{r sdo}
wss20 <- recodeList(wss20, list(
  list(column = "WSS10_1", recode_rules = five_r, new_column = "sdo_group_inferiority"),
  list(column = "WSS10_3", recode_rules = five_r, new_column = "sdo_antiequality_goal"),
  list(column = "WSS10_2", recode_rules = five_n, new_column = "sdo_unequal_deserving"),
  list(column = "WSS10_4", recode_rules = five_n, new_column = "sdo_unequal_chance")
))

sdo_items <- c(
  "sdo_group_inferiority", "sdo_antiequality_goal",
  "sdo_unequal_deserving", "sdo_unequal_chance"
)
psych::alpha(wss20[, sdo_items])
```

#### Racial Resentment (4 items, $\alpha$)

```{r racial-resentment}
wss20 <- recodeList(wss20, list(
  list(column = "WSS11_1", recode_rules = five_r, new_column = "rr_no_special_favors"),
  list(column = "WSS11_4", recode_rules = five_r, new_column = "rr_not_trying"),
  list(column = "WSS11_2", recode_rules = five_n, new_column = "rr_deny_slavery_legacy"),
  list(column = "WSS11_3", recode_rules = five_n, new_column = "rr_deny_underdeserving")
))

rr_items <- c(
  "rr_no_special_favors", "rr_not_trying",
  "rr_deny_slavery_legacy", "rr_deny_underdeserving"
)
psych::alpha(wss20[, rr_items])
```

#### Intergroup Empathy (4 items, $\alpha$)

```{r empathy}
wss20 <- recodeList(wss20, list(
  list(column = "WSS12_1", recode_rules = five_r, new_column = "empathy_tender_feelings"),
  list(column = "WSS12_2", recode_rules = five_r, new_column = "empathy_perspective_taking"),
  list(column = "WSS12_3", recode_rules = five_r, new_column = "empathy_imagine_before_criticize"),
  list(column = "WSS12_4", recode_rules = five_r, new_column = "empathy_protective")
))

empathy_items <- c(
  "empathy_tender_feelings", "empathy_perspective_taking",
  "empathy_imagine_before_criticize", "empathy_protective"
)
psych::alpha(wss20[, empathy_items])
```

#### Economic Individualism (6 items, $\alpha$)

```{r economic-individualism}
wss20 <- recodeList(wss20, list(
  list(column = "WSS13_1", recode_rules = five_r, new_column = "econ_indiv_hard_work_succeeds"),
  list(column = "WSS13_3", recode_rules = five_r, new_column = "econ_indiv_blame_self"),
  list(column = "WSS13_5", recode_rules = five_r, new_column = "econ_indiv_effort_rewarded"),
  list(column = "WSS13_2", recode_rules = five_n, new_column = "econ_indiv_deny_structural_barrier"),
  list(column = "WSS13_4", recode_rules = five_n, new_column = "econ_indiv_deny_ambition_futility"),
  list(column = "WSS13_6", recode_rules = five_n, new_column = "econ_indiv_deny_effort_futility")
))

econ_items <- c(
  "econ_indiv_hard_work_succeeds", "econ_indiv_blame_self",
  "econ_indiv_effort_rewarded", "econ_indiv_deny_structural_barrier",
  "econ_indiv_deny_ambition_futility", "econ_indiv_deny_effort_futility"
)
psych::alpha(wss20[, econ_items])
```

#### Affect / PANAS (6 items)

Negative affect (anxiety, anger, disgust) and positive affect (hope, pride,
enthusiasm) are separate sub-scales.

```{r affect}
wss20 <- recodeList(wss20, list(
  list(column = "WSS14_1", recode_rules = four_n, new_column = "affect_anxiety"),
  list(column = "WSS14_2", recode_rules = four_n, new_column = "affect_anger"),
  list(column = "WSS14_3", recode_rules = four_n, new_column = "affect_hope"),
  list(column = "WSS14_4", recode_rules = four_n, new_column = "affect_pride"),
  list(column = "WSS14_5", recode_rules = four_n, new_column = "affect_disgust"),
  list(column = "WSS14_6", recode_rules = four_n, new_column = "affect_enthusiasm")
))

neg_affect <- c("affect_anxiety", "affect_anger", "affect_disgust")
pos_affect <- c("affect_hope", "affect_pride", "affect_enthusiasm")
cat("Negative affect alpha:\n")
psych::alpha(wss20[, neg_affect])
cat("Positive affect alpha:\n")
psych::alpha(wss20[, pos_affect])
```

#### Institutional Trust (7 items, $\alpha$)

```{r trust}
wss20 <- recodeList(wss20, list(
  list(column = "WSS54_1", recode_rules = four_r, new_column = "trustCongress"),
  list(column = "WSS54_2", recode_rules = four_r, new_column = "trustPresident"),
  list(column = "WSS54_3", recode_rules = four_r, new_column = "trustSC"),
  list(column = "WSS54_4", recode_rules = four_r, new_column = "trustGovernment"),
  list(column = "WSS54_5", recode_rules = four_r, new_column = "trustStateleg"),
  list(column = "WSS54_6", recode_rules = four_r, new_column = "trustPolice"),
  list(column = "WSS54_7", recode_rules = four_r, new_column = "trustScience")
))

trust_items <- c(
  "trustCongress", "trustPresident", "trustSC",
  "trustGovernment", "trustStateleg", "trustPolice",
  "trustScience"
)
psych::alpha(wss20[, trust_items])
```

#### Policy Attitudes (WSS16: 12 items)

```{r policy}
wss20 <- recodeList(wss20, list(
  list(column = "WSS16_1", recode_rules = five_r, new_column = "policy_trade"),
  list(column = "WSS16_2", recode_rules = five_r, new_column = "policy_defense_spending"),
  list(column = "WSS16_3", recode_rules = five_r, new_column = "policy_budget_balance"),
  list(column = "WSS16_4", recode_rules = five_r, new_column = "policy_mandatory_minimums"),
  list(column = "WSS16_5", recode_rules = five_r, new_column = "policy_immigration"),
  list(column = "WSS16_6", recode_rules = five_r, new_column = "policy_paid_leave"),
  list(column = "WSS16_7", recode_rules = five_r, new_column = "policy_k12_education"),
  list(column = "WSS16_8", recode_rules = five_r, new_column = "policy_higher_education"),
  list(column = "WSS16_9", recode_rules = five_r, new_column = "policy_health"),
  list(column = "WSS16_10", recode_rules = five_r, new_column = "policy_marijuana"),
  list(column = "WSS16_11", recode_rules = five_r, new_column = "policy_social_media"),
  list(column = "WSS16_12", recode_rules = five_r, new_column = "policy_campaign_finance")
))
```

#### Rural-Urban Resentment (4 items, $\alpha$)

```{r rural-urban}
wss20 <- recodeList(wss20, list(
  list(column = "WSS31_1", recode_rules = five_r, new_column = "rural_urban_resentment_taxes"),
  list(column = "WSS31_2", recode_rules = five_r, new_column = "rural_urban_resentment_no_favors"),
  list(column = "WSS31_3", recode_rules = five_r, new_column = "rural_urban_resentment_move_away"),
  list(column = "WSS31_4", recode_rules = five_r, new_column = "rural_urban_resentment_too_much_say")
))

rur_items <- c(
  "rural_urban_resentment_taxes", "rural_urban_resentment_no_favors",
  "rural_urban_resentment_move_away", "rural_urban_resentment_too_much_say"
)
psych::alpha(wss20[, rur_items])
```

#### Environment (WSS17, WSS19, WSS20, WSS22)

```{r environment}
wss20 <- recodeList(wss20, list(
  list(column = "WSS17", recode_rules = five_n, new_column = "environmental_concern"),
  list(column = "WSS19_1", recode_rules = four_r, new_column = "envthreat_habitat_loss"),
  list(column = "WSS19_2", recode_rules = four_r, new_column = "envthreat_water_scarcity"),
  list(column = "WSS19_3", recode_rules = four_r, new_column = "envthreat_wildfires"),
  list(column = "WSS19_4", recode_rules = four_r, new_column = "envthreat_air_pollution"),
  list(column = "WSS19_5", recode_rules = four_r, new_column = "envthreat_rising_temps"),
  list(column = "WSS20_1", recode_rules = five_r, new_column = "envpolicy_water_infrastructure"),
  list(column = "WSS20_2", recode_rules = five_r, new_column = "envpolicy_water_conservation"),
  list(
    column = "WSS22", recode_rules = c(
      "1" = "energy_mining", "2" = "renewable_energy",
      "3" = "recreation", "4" = "preservation"
    ),
    new_column = "public_lands_priority"
  )
))

envthreat_items <- c(
  "envthreat_habitat_loss", "envthreat_water_scarcity",
  "envthreat_wildfires", "envthreat_air_pollution",
  "envthreat_rising_temps"
)
psych::alpha(wss20[, envthreat_items])
```

#### Immigration (WSS23, WSS61, WSS62)

```{r immigration}
wss20 <- recodeList(wss20, list(
  list(column = "WSS23_1", recode_rules = five_r, new_column = "immigration_legalization"),
  list(column = "WSS23_2", recode_rules = five_r, new_column = "immigration_end_separation"),
  list(column = "WSS23_3", recode_rules = five_r, new_column = "immigration_reduce_asylum"),
  list(column = "WSS23_4", recode_rules = five_r, new_column = "immigration_end_birthright"),
  list(column = "WSS23_5", recode_rules = five_r, new_column = "immigration_border_tech"),
  list(column = "WSS23_6", recode_rules = five_r, new_column = "immigration_border_wall"),
  list(column = "WSS23_7", recode_rules = five_r, new_column = "immigration_legal_resources"),
  list(column = "WSS61_1", recode_rules = c("1" = 1, "2" = 0), new_column = "know_undocumented"),
  list(column = "WSS61_2", recode_rules = c("1" = 1, "2" = 0), new_column = "know_daca"),
  list(column = "WSS62", recode_rules = five_r, new_column = "immigration_deportation_worry")
))

immig_items <- c(
  "immigration_legalization", "immigration_end_separation",
  "immigration_reduce_asylum", "immigration_end_birthright"
)
psych::alpha(wss20[, immig_items])
```

#### Law Enforcement and Firearms

```{r law-firearms}
wss20 <- recodeList(wss20, list(
  list(column = "WSS25", recode_rules = five_n, new_column = "law_enforcement_rating"),
  list(column = "WSS26", recode_rules = c("1" = 1, "2" = 0), new_column = "firearm_in_home"),
  list(column = "WSS27_1", recode_rules = five_r, new_column = "firearms_concealed_carry"),
  list(column = "WSS27_2", recode_rules = five_r, new_column = "firearms_background_checks"),
  list(column = "WSS27_3", recode_rules = five_r, new_column = "firearms_national_registry")
))

firearms_items <- c(
  "firearms_concealed_carry", "firearms_background_checks",
  "firearms_national_registry"
)
psych::alpha(wss20[, firearms_items])
```

#### COVID (WSS47, WSS48)

```{r covid}
wss20 <- recodeList(wss20, list(
  list(column = "WSS47_1", recode_rules = five_r, new_column = "covid_mask_mandate"),
  list(column = "WSS48_1", recode_rules = binary_10, new_column = "covid_social_distance"),
  list(column = "WSS48_2", recode_rules = binary_10, new_column = "covid_quarantine"),
  list(column = "WSS48_3", recode_rules = binary_10, new_column = "covid_check_neighbors"),
  list(column = "WSS48_4", recode_rules = binary_10, new_column = "covid_wore_mask"),
  list(column = "WSS48_5", recode_rules = binary_10, new_column = "covid_protest_racial"),
  list(column = "WSS48_6", recode_rules = binary_10, new_column = "covid_protest_guidelines")
))

covid_compliance <- c(
  "covid_social_distance", "covid_quarantine",
  "covid_check_neighbors", "covid_wore_mask"
)
psych::alpha(wss20[, covid_compliance])
```

#### Political Participation (9 items, $\alpha$) and Consumerism

```{r participation}
wss20 <- recodeList(wss20, list(
  list(column = "WSS32_1", recode_rules = binary_10, new_column = "participation_rally"),
  list(column = "WSS32_2", recode_rules = binary_10, new_column = "participation_sign"),
  list(column = "WSS32_3", recode_rules = binary_10, new_column = "participation_volunteer"),
  list(column = "WSS32_4", recode_rules = binary_10, new_column = "participation_protest"),
  list(column = "WSS32_5", recode_rules = binary_10, new_column = "participation_contact_official"),
  list(column = "WSS32_6", recode_rules = binary_10, new_column = "participation_donate"),
  list(column = "WSS32_7", recode_rules = binary_10, new_column = "participation_social_media"),
  list(column = "WSS32_8", recode_rules = binary_10, new_column = "participation_persuade"),
  list(column = "WSS32_9", recode_rules = binary_10, new_column = "participation_none"),
  list(column = "WSS34_a", recode_rules = binary_10, new_column = "consumerism_buycott"),
  list(column = "WSS34_b", recode_rules = binary_10, new_column = "consumerism_political_purchase")
))

part_items <- c(
  "participation_rally", "participation_sign",
  "participation_volunteer", "participation_protest",
  "participation_contact_official", "participation_donate",
  "participation_social_media"
)
psych::alpha(wss20[, part_items])
```

#### Efficacy, Census, and Election Integrity

```{r efficacy-election}
wss20 <- recodeList(wss20, list(
  list(column = "WSS33_1", recode_rules = five_n, new_column = "internal_efficacy"),
  list(column = "WSS33_2", recode_rules = five_n, new_column = "external_efficacy"),
  list(column = "WSS35_b", recode_rules = four_r, new_column = "census_confidence"),
  list(column = "WSS41_6", recode_rules = five_r, new_column = "election_concern_delayed_results"),
  list(column = "WSS41_7", recode_rules = five_r, new_column = "election_concern_ineligible_voting")
))
```

#### News Media (11 sources)

```{r news-media}
wss20 <- recodeList(wss20, list(
  list(column = "WSS58_1", recode_rules = binary_10, new_column = "news_cnn"),
  list(column = "WSS58_2", recode_rules = binary_10, new_column = "news_network_tv"),
  list(column = "WSS58_3", recode_rules = binary_10, new_column = "news_national_paper"),
  list(column = "WSS58_4", recode_rules = binary_10, new_column = "news_fox"),
  list(column = "WSS58_5", recode_rules = binary_10, new_column = "news_msnbc"),
  list(column = "WSS58_6", recode_rules = binary_10, new_column = "news_local_broadcast"),
  list(column = "WSS58_7", recode_rules = binary_10, new_column = "news_local_paper"),
  list(column = "WSS58_8", recode_rules = binary_10, new_column = "news_social_media"),
  list(column = "WSS58_9", recode_rules = binary_10, new_column = "news_npr_pbs"),
  list(column = "WSS58_10", recode_rules = binary_10, new_column = "news_telemundo"),
  list(column = "WSS58_95", recode_rules = binary_10, new_column = "news_other")
))
```

#### Income Gap and Social Mobility

```{r income-mobility}
wss20 <- recodeList(wss20, list(
  list(column = "WSS59a", recode_rules = c(
    "1" = "background_determines", "2" = "fair_chance",
    "3" = "neither"
  ), new_column = "income_gap_belief"),
  list(column = "WSS60a_1", recode_rules = three_n, new_column = "redist_less_well_off"),
  list(column = "WSS60a_2", recode_rules = three_n, new_column = "redist_just_managing"),
  list(column = "WSS60a_3", recode_rules = three_n, new_column = "redist_comfortable"),
  list(column = "WSS60a_4", recode_rules = three_n, new_column = "redist_most_well_off")
))

redist_items <- c(
  "redist_less_well_off", "redist_just_managing",
  "redist_comfortable", "redist_most_well_off"
)
psych::alpha(wss20[, redist_items])
```

#### Police/Crime Contact (WSS63m)

```{r police-contact}
wss20 <- recodeList(wss20, list(
  list(column = "WSS63m_1_1", recode_rules = binary_10, new_column = "police_stopped_self"),
  list(column = "WSS63m_1_2", recode_rules = binary_10, new_column = "police_stopped_other"),
  list(column = "WSS63m_1_3", recode_rules = binary_10, new_column = "police_stopped_none"),
  list(column = "WSS63m_2_1", recode_rules = binary_10, new_column = "arrested_self"),
  list(column = "WSS63m_2_2", recode_rules = binary_10, new_column = "arrested_other"),
  list(column = "WSS63m_2_3", recode_rules = binary_10, new_column = "arrested_none"),
  list(column = "WSS63m_3_1", recode_rules = binary_10, new_column = "crime_victim_self"),
  list(column = "WSS63m_3_2", recode_rules = binary_10, new_column = "crime_victim_other"),
  list(column = "WSS63m_3_3", recode_rules = binary_10, new_column = "crime_victim_none"),
  list(column = "WSS63m_4_1", recode_rules = binary_10, new_column = "called_police_self"),
  list(column = "WSS63m_4_2", recode_rules = binary_10, new_column = "called_police_other"),
  list(column = "WSS63m_4_3", recode_rules = binary_10, new_column = "called_police_none")
))
```

#### Vote Choice and Electoral Contestation (5 items, $\alpha$)

```{r vote-contestation}
wss20 <- recodeList(wss20, list(
  list(column = "WSS36_b", recode_rules = binary_10, new_column = "v1"),
  list(column = "WSS36_c", recode_rules = binary_10, new_column = "v2"),
  list(column = "WSS36_a", recode_rules = c("1" = 1, "2" = 0, "4" = 0), new_column = "voted_2020"),
  list(column = "WSS40_1", recode_rules = five_r, new_column = "tmp_contestation_attend_march"),
  list(column = "WSS40_2", recode_rules = five_r, new_column = "tmp_contestation_criticize_election"),
  list(column = "WSS40_3", recode_rules = five_r, new_column = "tmp_contestation_burn_flag"),
  list(column = "WSS40_4", recode_rules = five_r, new_column = "tmp_contestation_recount"),
  list(column = "WSS40_5", recode_rules = five_r, new_column = "tmp_contestation_court")
))

contest_items <- c(
  "tmp_contestation_attend_march", "tmp_contestation_criticize_election",
  "tmp_contestation_burn_flag", "tmp_contestation_recount",
  "tmp_contestation_court"
)
psych::alpha(wss20[, contest_items])
```

#### Arizona State Items

```{r arizona}
therm_11pt <- c(
  "1" = 0, "2" = 1, "3" = 2, "4" = 3, "5" = 4,
  "6" = 5, "7" = 6, "8" = 7, "9" = 8, "10" = 9, "11" = 10
)

wss20 <- recodeList(wss20, list(
  list(column = "AZ301", recode_rules = therm_11pt, new_column = "az_therm_dem"),
  list(column = "AZ302", recode_rules = therm_11pt, new_column = "az_therm_rep"),
  list(column = "AZ303", recode_rules = therm_11pt, new_column = "az_therm_ind"),
  list(column = "AZ304", recode_rules = therm_11pt, new_column = "az_hisp_therm_dem"),
  list(column = "AZ305", recode_rules = therm_11pt, new_column = "az_hisp_therm_rep"),
  list(column = "AZ306", recode_rules = therm_11pt, new_column = "az_hisp_therm_ind"),
  list(column = "AZ307", recode_rules = five_r, new_column = "identity_racial_importance"),
  list(column = "AZ308", recode_rules = five_r, new_column = "identity_state_importance"),
  list(column = "AZ309", recode_rules = five_r, new_column = "identity_partisan_importance"),
  list(column = "AZ311_pipe_name", recode_rules = c("1" = "McSally", "2" = "Kelly"), new_column = "az_senate_candidate"),
  list(
    column = "AZ311", recode_rules = c(
      "1" = "strong_pro_candidate", "2" = "weak_pro_candidate",
      "3" = "weak_anti_opponent", "4" = "strong_anti_opponent"
    ),
    new_column = "az_senate_justification"
  )
))
```


### Scale Construction

Scales are created only when (a) all items share the same response metric
and (b) Cronbach's $\alpha \geq 0.60$. Each scale is the `rowMeans()` of its
items, rescaled to 0--1 via `zero_one()`.

```{r scale-helper}
# Helper: compute alpha, build scale if >= threshold
make_scale <- function(data, items, scale_name, threshold = 0.60) {
  mat <- data[, items, drop = FALSE]
  a <- psych::alpha(mat, check.keys = TRUE)
  cat(sprintf(
    "%s: alpha = %.3f (n items = %d)\n",
    scale_name, a$total$raw_alpha, length(items)
  ))
  if (a$total$raw_alpha >= threshold) {
    data[[scale_name]] <- zero_one(rowMeans(mat, na.rm = TRUE))
    cat("  -> Scale created.\n")
  } else {
    cat("  -> Alpha below threshold; scale NOT created.\n")
  }
  data
}
```

```{r build-scales}
# ── Authoritarianism (4 binary items, same metric) ──
wss20 <- make_scale(
  wss20,
  c("auth_1", "auth_2", "auth_3", "auth_4"),
  "authoritarianism"
)

# ── Social Dominance Orientation (4 items, all 5-pt) ──
wss20 <- make_scale(
  wss20,
  c(
    "sdo_group_inferiority", "sdo_antiequality_goal",
    "sdo_unequal_deserving", "sdo_unequal_chance"
  ),
  "sdo_index"
)

# ── Racial Resentment (4 items, all 5-pt) ──
wss20 <- make_scale(
  wss20,
  c(
    "rr_no_special_favors", "rr_not_trying",
    "rr_deny_slavery_legacy", "rr_deny_underdeserving"
  ),
  "racial_resentment_index"
)

# ── Intergroup Empathy (4 items, all 5-pt) ──
wss20 <- make_scale(
  wss20,
  c(
    "empathy_tender_feelings", "empathy_perspective_taking",
    "empathy_imagine_before_criticize", "empathy_protective"
  ),
  "intergroup_empathy_index"
)

# ── Economic Individualism (6 items, all 5-pt) ──
wss20 <- make_scale(
  wss20,
  c(
    "econ_indiv_hard_work_succeeds", "econ_indiv_blame_self",
    "econ_indiv_effort_rewarded", "econ_indiv_deny_structural_barrier",
    "econ_indiv_deny_ambition_futility", "econ_indiv_deny_effort_futility"
  ),
  "economic_individualism_index"
)

# ── Negative Affect (3 items, all 4-pt) ──
wss20 <- make_scale(
  wss20,
  c("affect_anxiety", "affect_anger", "affect_disgust"),
  "negative_affect"
)

# ── Positive Affect (3 items, all 4-pt) ──
wss20 <- make_scale(
  wss20,
  c("affect_hope", "affect_pride", "affect_enthusiasm"),
  "positive_affect"
)

# ── Institutional Trust (7 items, all 4-pt) ──
wss20 <- make_scale(
  wss20,
  c(
    "trustCongress", "trustPresident", "trustSC",
    "trustGovernment", "trustStateleg", "trustPolice",
    "trustScience"
  ),
  "trust_index"
)

# ── Rural-Urban Resentment (4 items, all 5-pt) ──
wss20 <- make_scale(
  wss20,
  c(
    "rural_urban_resentment_taxes", "rural_urban_resentment_no_favors",
    "rural_urban_resentment_move_away", "rural_urban_resentment_too_much_say"
  ),
  "rural_urban_resentment_index"
)

# ── Environmental Threat (5 items, all 4-pt) ──
wss20 <- make_scale(
  wss20,
  c(
    "envthreat_habitat_loss", "envthreat_water_scarcity",
    "envthreat_wildfires", "envthreat_air_pollution",
    "envthreat_rising_temps"
  ),
  "envthreat_index"
)

# ── Immigration Attitudes (4 core items, all 5-pt) ──
# Note: WSS23_5–7 are asked of a subsample; using the 4 universal items
wss20 <- make_scale(
  wss20,
  c(
    "immigration_legalization", "immigration_end_separation",
    "immigration_reduce_asylum", "immigration_end_birthright"
  ),
  "immigration_index"
)

# ── Firearms Policy (3 items, all 5-pt) ──
wss20 <- make_scale(
  wss20,
  c(
    "firearms_concealed_carry", "firearms_background_checks",
    "firearms_national_registry"
  ),
  "firearms_index"
)

# ── COVID Compliance (4 items, all binary) ──
wss20 <- make_scale(
  wss20,
  c(
    "covid_social_distance", "covid_quarantine",
    "covid_check_neighbors", "covid_wore_mask"
  ),
  "covid_compliance_index"
)

# ── Political Participation (7 items, all binary) ──
# Excludes participation_persuade and participation_none
wss20 <- make_scale(
  wss20,
  c(
    "participation_rally", "participation_sign",
    "participation_volunteer", "participation_protest",
    "participation_contact_official", "participation_donate",
    "participation_social_media"
  ),
  "political_participation_index"
)

# ── Redistribution (4 items, all 3-pt) ──
wss20 <- make_scale(
  wss20,
  c(
    "redist_less_well_off", "redist_just_managing",
    "redist_comfortable", "redist_most_well_off"
  ),
  "redistribution_index"
)

# ── Electoral Contestation (5 items, all 5-pt) ──
wss20 <- make_scale(
  wss20,
  c(
    "tmp_contestation_attend_march", "tmp_contestation_criticize_election",
    "tmp_contestation_burn_flag", "tmp_contestation_recount",
    "tmp_contestation_court"
  ),
  "contestation_index"
)

# ── NV Intergenerational Mobility (4 items, all 3-pt) ──
wss20 <- make_scale(
  wss20,
  c(
    "nv_intergen_education", "nv_intergen_living_standard",
    "nv_intergen_housing", "nv_intergen_social_position"
  ),
  "nv_intergenerational_index"
)
```

### Derived Variables

```{r derived-variables}
wss20 <- wss20 %>%
  mutate(
    presvote_trump_2020 = ifelse(WSS36_b == 9, v2, v1),
    vote_trump = presvote_trump_2020,
    republican = ifelse(party_identification3 == 3, 1, 0),
    democrat = ifelse(party_identification3 == 1, 1, 0),
    independent = ifelse(party_identification3 == 2, 1, 0),
    libcon = (ideology - 1) / 4,
    prepost = ifelse(prepost == "post", 1, 0)
  )

# Trim to recoded columns only
overwritten <- c(
  "age", "state", "female", "black", "white", "latino",
  "asian", "american_indian", "married", "college", "faminc"
)
new_columns <- setdiff(names(wss20), original_columns)
new_columns <- unique(c(new_columns, overwritten, "date", "staticDate"))
wss20 <- wss20 %>% select(all_of(new_columns))
```

### Propensity Scores

```{r}
ps_out_20 <- ps_weight(wss20, "prepost", ~ female + latino +
  white + black + asian + college + age + vote_trump + as.factor(state), methods = "glm")
wss20 <- ps_out_20$data
```

### Reshape to Long Format

```{r}
wss20 <- wss20 %>%
  pivot_longer(
    cols = starts_with("tmp_contestation_"),
    names_to = "contestation",
    values_to = "contestation_value",
    values_drop_na = FALSE
  ) %>%
  mutate(
    survey_wave = ifelse(prepost == 1, "post", "pre"),
    contestation = str_remove(contestation, "^tmp_contestation_")
  ) %>%
  arrange(caseidID) |>
  group_by(caseidID) |>
  mutate(
    electoral_contestation = mean(contestation_value, na.rm = TRUE)
  ) |>
  ungroup()

wss20$metadata <- "2020 WSS: natural experiment pre-post design."
```
