---
title: "Getting Started with vrm"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with vrm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4
)
```

## Overview

`vrm`, **Visualization and Regression Modeling**, is an R package designed to streamline the workflow of regression modeling, prediction, and visualization for social science research. It provides a unified interface for regression modeling, posterior prediction, counterfactual inference, and publication-quality visualization. This vignette demonstrates the workflow using the package's built-in survey data.

## 1. Data Utilities

### Rescaling

Use `zero_one()` or `zero.one()` to rescale a numeric variable to the [0, 1] range:

```{r rescale}
library(vrm)

# Rescale a simple vector
x <- c(10, 20, 30, 40, 50)
zero_one(x)

# Works with NAs
zero_one(c(1, NA, 5, 10), na.rm = TRUE)
```

### Survey Recoding

The package provides named recode vectors for common Likert scales:

```{r recode-vectors}
# Reverse-code a 5-point scale
five_mn

# Normal 4-point scale
four_n
```

Use `recodeColumn()` for a single column, or `recodeList()` to recode many columns at once:

```{r recode-example}
# Create sample data mimicking raw survey responses
sample_df <- data.frame(
  id = 1:5,
  q1_raw = c(1, 2, 3, 4, 5),
  q2_raw = c(4, 3, 2, 1, 4)
)

# Recode a single column (reverse 5-point)
sample_df <- recodeColumn(sample_df,
  column = "q1_raw",
  recode_rules = five_r, colname = "q1"
)
sample_df[, c("q1_raw", "q1")]

rules <- list(
  list(column = "q1_raw", recode_rules = five_r, new_column = "q1_rev"),
  list(column = "q2_raw", recode_rules = four_r, new_column = "q2_rev")
)
sample_df <- recodeList(sample_df, rules)
sample_df
```

The function uses `dplyr::recode()` under the hood, but minimizes the need to write out the recoding rules manually using `case_when()`. 


## 2. Exploratory Data Analysis

For simple data checking and verification, use the EDA helpers. `summarize_missingness()` displays missing data patterns,
`describe_numeric()` displays descriptive statistics for a numeric variable, `variable_characteristics()` displays variable characteristics, and `cor_table()` displays correlation tables.

```{r eda}
# Create example data
set.seed(42)
demo_df <- data.frame(
  age = rnorm(200, 45, 15),
  income = sample(c(1:5, NA), 200, replace = TRUE),
  education = sample(c("HS", "BA", "MA", NA), 200, replace = TRUE),
  score = rnorm(200, 50, 10)
)

# Summarize missingness
summarize_missingness(demo_df)

# Descriptive statistics for a numeric variable
describe_numeric(demo_df$age)

# Variable characteristics overview
variable_characteristics(demo_df)

# Correlation table (long format, for numeric columns)
cor_tbl <- cor_table(demo_df[, c("age", "score")])
head(cor_tbl)
```

## 3. Ordered Regression (Frequentist)

`R` has numerous packages for fitting regression models. The goal is not to reinvent the wheel, but instead write  
wrappers around existing packages to provide a unified interface.  First, ordered logit and probit models using the `MASS::polr()` wrappers:

```{r ordered-models}

# Create ordered factor data
set.seed(123)
n <- 200
x1 <- rnorm(n)
x2 <- rnorm(n)
latent <- 0.8 * x1 - 0.5 * x2 + rnorm(n)
y <- cut(latent,
  breaks = c(-Inf, -1, 0, 1, Inf),
  labels = c("Poor", "Fair", "Good", "Excellent")
)
ord_df <- data.frame(y = y, x1 = x1, x2 = x2)

# Fit ordered logit
model_ologit <- fit_ordered_logit(y ~ x1 + x2, data = ord_df)

# Predict classes
preds <- predict_ordered(model_ologit, type = "class")
table(preds)

# Predicted probabilities for new data
new_data <- data.frame(x1 = c(-1, 0, 1), x2 = c(0, 0, 0))
predict_ordered(model_ologit, newdata = new_data, type = "probs")

# Enhanced summary with p-values
options(scipen = 6)  # Favor fixed/decimal notation
model_summary <- summary_ordered(model_ologit)
model_summary$coefficients
```

## 4. Frequentist Predicted Probabilities with Uncertainty


Generating predictions from ordered logit models is more annoying than it should be, especially if you want to incorporate uncertainty. The `predict_ordinal_probs()` function generates predicted probabilities for each category, along with uncertainty estimates based on simulation draws from the model's parameter distributions.

### Ordered Model Predictions

```{r predict-ordinal}
# Create a design matrix of scenarios
design <- expand.grid(x1 = seq(-2, 2, by = 1), x2 = 0)

# Predict category probabilities with uncertainty (1000 draws)
pred_ord <- predict_ordinal_probs(design, model_ologit, n_draws = 1000)
head(pred_ord)

# Summarize across draws
summary_ord <- summarize_predictions(pred_ord, prob_col = "probability")
# Favor fixed/decimal notation for summary output
options(scipen = 6)
summary_ord
```

### Binary Logit Predictions

```{r predict-logit}
# Fit a binary logistic regression
bin_df <- data.frame(
  outcome = rbinom(300, 1, plogis(0.5 * x1[1:300 %% n + 1])),
  x1 = x1[1:300 %% n + 1],
  x2 = x2[1:300 %% n + 1]
)
model_logit <- glm(outcome ~ x1 + x2, data = bin_df, family = binomial)

# Predicted probabilities with uncertainty
design_bin <- data.frame(x1 = seq(-2, 2, 0.5), x2 = 0)
pred_logit <- predict_logit_probs(design_bin, model_logit, n_draws = 500)
summary_logit <- summarize_predictions(pred_logit, prob_col = "predicted_prob")
summary_logit
```

## 5. Counterfactual Analysis (G-Computation)

Create counterfactual datasets for g-computation style causal inference:

```{r counterfactual}
# Sample data with a binary treatment
cf_df <- data.frame(
  id = 1:6,
  treatment = c(0, 0, 0, 1, 1, 1),
  age = c(25, 35, 45, 30, 40, 50),
  outcome = c(3, 4, 3, 5, 4, 5)
)

# Generate counterfactual data (flips treatment 0↔1)
full_set <- generate_counterfactual_data(cf_df, treatment_var = "treatment")
full_set[, c("id", "treatment", "original_treatment", "pre_condition")]
```

The `full_set` now has doubled rows: original observations tagged `"actual_treatment"`
and counterfactual versions tagged `"hypothetical_treatment"`. This is designed to feed
into `brms::add_epred_draws()` for posterior counterfactual predictions.

## 6. Propensity Score Weighting

Estimate inverse probability of treatment weights (IPTW) using logistic regression:

```{r ps-weight}
# Create treatment data
set.seed(99)
n_ps <- 500
ps_df <- data.frame(
  treatment = rbinom(n_ps, 1, 0.4),
  age = rnorm(n_ps, 40, 10),
  female = rbinom(n_ps, 1, 0.5),
  college = rbinom(n_ps, 1, 0.3)
)

# Estimate propensity scores and IPTW weights (GLM only)
ps_out <- ps_weight(ps_df,
  treatment = "treatment",
  formula = ~ age + female + college,
  methods = "glm"
)

# The weighted data includes ps and wts columns
head(ps_out$data[, c("treatment", "age", "ps", "wts")])

# Summary of weight distribution
summary(ps_out$data$wts)
```

## 7. Visualization Theme

Apply the `vrm` publication theme to any ggplot:
```{r theme-demo}
library(ggplot2)

p <- ggplot(
  data.frame(x = rnorm(200), y = rnorm(200)),
  aes(x, y)
) +
  geom_point(alpha = 0.5) +
  labs(title = "Example with mytheme()", x = "X Variable", y = "Y Variable") +
  mytheme()
p
```

## 8. Bayesian Workflow (brms)

The following demonstrates the Bayesian workflow using `brms`. These chunks
are not evaluated here because `brms` requires Stan compilation, but they
show the intended usage.

```{r bayesian-workflow, eval = FALSE}
library(brms)
library(tidybayes)

# Load package data
data(wss20, package = "vrm")

# Rescale age
wss20$age_scaled <- zero_one(wss20$age)

# Fit a Bayesian ordinal model
fit_ord <- brms.ordinal(
  data = wss20,
  DV = "contestation_value",
  IV = "prepost * vote_trump",
  cores = 4, chains = 4, iter = 2000
)

# Or use the general-purpose brms_fit()
fit_lin <- brms_fit(
  data = wss20,
  DV = "contestation_value",
  IV = "prepost * vote_trump"
)

# Posterior predicted means
preds <- posterior_means(fit_lin, xvar = "prepost", mvar = "vote_trump")

# Posterior marginal effects
me <- posterior_pme(fit_lin, xvar = "prepost", mvar = "vote_trump")

# Counterfactual g-computation
cf_data <- generate_counterfactual_data(wss20, treatment_var = "prepost")
preds_long <- cf_data %>%
  tidybayes::add_epred_draws(fit_lin, ndraws = 500)
ind_effects <- calculate_individual_effects(
  preds_long,
  treatment_var = "prepost",
  group_vars = c("vote_trump")
)

# Visualization
ggMargins_categories(fit_ord, data = wss20)
gg_posterior_ridges(fit_lin, data = wss20)
```

## Summary

The typical `vrm` workflow is:

1. **Recode** survey data with `recodeList()` / `zero_one()`
2. **Explore** with `summarize_missingness()`, `describe_numeric()`, `cor_table()`
3. **Weight** (if observational) with `ps_weight()`
4. **Fit** ordered models with `fit_ordered_logit()` or Bayesian models with `brms.ordinal()` / `brms_fit()`
5. **Predict** with `predict_ordinal_probs()` / `predict_logit_probs()` or `posterior_means()` / `posterior_pme()`
6. **Counterfactual** analysis with `generate_counterfactual_data()` + `calculate_individual_effects()`
7. **Visualize** with `plotSunflower()`, `gg_posterior_ridges()`, `mytheme()`, etc.
