---
title: "Getting Started with vrm"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with vrm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

## Overview

`vrm` provides a unified interface for regression modeling, posterior prediction,
counterfactual inference, and publication-quality visualization. This vignette
walks through the core workflow using electoral contestation survey data.

## Data Preparation

```{r}
library(vrm)
library(dplyr)

# Load example data (included with package)
data(wss20, package = "vrm")

# Rescale age to [0,1]
wss20$age_scaled <- zero_one(wss20$age)

# Bulk recode items using recode rules
rules <- list(
  list(column = "raw_item1", recode_rules = five_r, new_column = "item1"),
  list(column = "raw_item2", recode_rules = four_r, new_column = "item2")
)
wss20 <- recodeList(wss20, rules)
```

## Propensity Score Weighting

For observational pre/post designs, use `ps_weight()` to estimate inverse
probability of treatment weights (IPTW):

```{r}
ps_out <- ps_weight(
  wss20,
  treatment = "prepost",
  formula = ~ female + latino + white + black + college + age_scaled,
  methods = "glm"
)

# The weighted data is in ps_out$data
wss20 <- ps_out$data
```

## Model Fitting

Fit a Bayesian ordinal model with a pre/post × vote interaction:

```{r}
fit_ord <- brms.ordinal(
  dv = "contestation_value",
  iv = "prepost * vote_trump",
  data = wss20,
  cores = 4,
  chains = 4,
  iter = 2000
)
```

Or a linear model:

```{r}
fit_lin <- brms.linear(
  dv = "contestation_value",
  iv = "prepost * vote_trump",
  data = wss20,
  cores = 4
)
```

## Posterior Predictions

Generate predicted means across conditions:

```{r}
preds <- posterior_means(
  fit_lin,
  iv_name = "prepost",
  moderator = "vote_trump",
  data = wss20
)
```

## Counterfactual Analysis (G-Computation)

Create counterfactual data and compute individual treatment effects:

```{r}
# Generate observed + counterfactual rows
cf_data <- generate_counterfactual_data(wss20, treatment_var = "prepost")

# Get posterior predictions for both conditions
library(tidybayes)
preds_long <- cf_data %>%
  add_epred_draws(fit_lin, ndraws = 500)

# Calculate individual-level effects
ind_effects <- calculate_individual_effects(
  preds_long,
  treatment_var = "prepost",
  group_vars = c("vote_trump")
)
```

## Visualization

```{r}
library(ggplot2)

# Marginal effects for ordinal model
ggMargins_categories(fit_ord, data = wss20)

# Ridgeline plot of posterior draws
gg_posterior_ridges(fit_lin, data = wss20)
```

## Summary

The typical `vrm` workflow is:

1. **Recode** survey data with `recodeList()` / `zero_one()`
2. **Weight** (if observational) with `ps_weight()`
3. **Fit** models with `brms.ordinal()`, `brms.linear()`, etc.
4. **Predict** with `posterior_means()` / `posterior_pme()`
5. **Counterfactual** analysis with `generate_counterfactual_data()` + `calculate_individual_effects()`
6. **Visualize** with `ggMargins_categories()`, sunflower plots, ridgeplots, etc.
