---
title: "Visualization Gallery"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualization Gallery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  eval = FALSE
)
```

## Overview

This vignette demonstrates the visualization tools in `vrm` for communicating Bayesian estimation results. Each plot type serves a distinct purpose in the post-estimation workflow: ridgeplots show full posterior distributions, split violin plots compare distributions across conditions, forest plots summarize point estimates with uncertainty, sunflower plots display predicted probabilities alongside observed data, and combined panels bring these together for publication.

All examples assume fitted `brms` models from the Western States Survey (`wss20`). See `vignette("bayesian-estimation")` for the estimation workflow.

```{r libraries}
library(vrm)
library(brms)
library(tidybayes)
library(dplyr)
library(tidyr)
library(ggplot2)
library(purrr)
library(ggridges)
library(patchwork)

data(wss20, package = "vrm")
```

## Assumed Models

The examples below reference models fitted in the Bayesian estimation vignette. Here is the setup code (not re-run):

```{r fit-models}
items <- c("attend_march", "burn_flag", "recount", "court", "criticize_election")
families <- list(ord = cumulative(), linear = gaussian())

models <- fit_brms_battery(
  data     = wss20,
  items    = items,
  rhs      = "prepost * vote_trump + female + latino + college + scale(age)",
  families = families,
  data_dir = tempdir(),
  prefix   = "wss20",
  weight_var = NULL,
  cores = 4, chains = 3, iter = 2000
)
```

## 1. Ridgeplots: Full Posterior Distributions

Ridgeplots (joy plots) display the entire posterior predictive distribution for each item, filled by condition and faceted by moderator. They reveal distributional differences that point summaries miss.

```{r ridgeplot}
# Collect posterior draws across items
all_draws <- map_dfr(items, function(it) {
  posterior_draws_grid(
    model  = models[[paste0(it, "_linear")]],
    xvar   = "prepost",
    mvar   = "vote_trump",
    xval   = c(0, 1),
    mval   = c(0, 1),
    ndraws = 500
  ) %>%
    mutate(item = it)
})

# Basic ridgeplot
gg_posterior_ridges(
  draws        = all_draws,
  ridge_var    = "item",
  fill_var     = "prepost",
  facet_var    = "vote_trump",
  fill_labels  = c("0" = "Pre-Election", "1" = "Post-Election"),
  facet_labels = c("0" = "Biden Voter", "1" = "Trump Voter"),
  ridge_labels = c(
    "attend_march"       = "Attend March",
    "burn_flag"          = "Burn Flag",
    "recount"            = "Recount",
    "court"              = "Go to Court",
    "criticize_election" = "Criticize Election"
  ),
  title = "Posterior Predicted Mean Support",
  xlab  = "Predicted Mean \u2192"
)
```

### Customizing the ridgeplot

```{r ridgeplot-custom}
# Different palette, no quantile lines
gg_posterior_ridges(
  draws          = all_draws,
  ridge_var      = "item",
  fill_var       = "prepost",
  facet_var      = "vote_trump",
  viridis_option = "mako",
  alpha          = 0.6,
  scale          = 1.2,
  quantile_lines = FALSE,
  fill_labels    = c("0" = "Pre", "1" = "Post"),
  facet_labels   = c("0" = "Biden", "1" = "Trump"),
  title          = "Predicted Support (Mako palette)"
)
```

### Ridgeplot for ordinal models

For ordinal models, filter to a specific response category (e.g., Pr(Strongly Support)):

```{r ridgeplot-ordinal}
ord_draws <- map_dfr(items, function(it) {
  posterior_draws_grid(
    model  = models[[paste0(it, "_ord")]],
    xvar   = "prepost",
    mvar   = "vote_trump",
    xval   = c(0, 1),
    mval   = c(0, 1),
    ndraws = 500
  ) %>%
    mutate(item = it)
})

gg_posterior_ridges(
  draws    = ord_draws,
  ridge_var = "item",
  fill_var  = "prepost",
  facet_var = "vote_trump",
  category  = "5",
  fill_labels  = c("0" = "Pre-Election", "1" = "Post-Election"),
  facet_labels = c("0" = "Biden Voter", "1" = "Trump Voter"),
  title = "Pr(Strongly Support)"
)
```

## 2. Split Violin Plots

Split violin plots are ideal for comparing two posterior distributions side by side. We build these directly from posterior draws using `ggplot2`. The split violin geometry places pre- and post-election distributions on opposite sides of the same axis.

```{r split-violin-geom}
# Custom GeomSplitViolin for ggplot2
GeomSplitViolin <- ggproto("GeomSplitViolin", GeomViolin,
  draw_group = function(self, data, ..., draw_quantiles = NULL) {
    data <- transform(data,
      xminv = x - violinwidth * (x - xmin),
      xmaxv = x + violinwidth * (xmax - x)
    )
    grp <- data[1, "group"]
    newdata <- plyr::arrange(
      transform(data, x = if (grp %% 2 == 1) xminv else xmaxv),
      if (grp %% 2 == 1) y else -y
    )
    newdata <- rbind(newdata[1, ], newdata, newdata[nrow(newdata), ], newdata[1, ])
    newdata[c(1, nrow(newdata) - 1, nrow(newdata)), "x"] <- round(newdata[1, "x"], 5)
    ggplot2:::ggname("geom_split_violin",
      grid::grobTree(GeomPolygon$draw_panel(newdata, ...))
    )
  }
)

geom_split_violin <- function(mapping = NULL, data = NULL, stat = "ydensity",
                               position = "identity", ..., draw_quantiles = NULL,
                               trim = TRUE, scale = "area", na.rm = FALSE,
                               show.legend = NA, inherit.aes = TRUE) {
  layer(
    data = data, mapping = mapping, stat = stat, geom = GeomSplitViolin,
    position = position, show.legend = show.legend, inherit.aes = inherit.aes,
    params = list(trim = trim, scale = scale, draw_quantiles = draw_quantiles,
                  na.rm = na.rm, ...)
  )
}
```

### Split violin by pre/post condition

```{r split-violin-plot}
# Use draws from a single item
draws_recount <- posterior_draws_grid(
  model  = models[["recount_linear"]],
  xvar   = "prepost",
  mvar   = "vote_trump",
  xval   = c(0, 1),
  mval   = c(0, 1),
  ndraws = 500
)

draws_recount %>%
  ungroup() %>%
  mutate(
    vote_trump = factor(vote_trump, levels = c(0, 1),
                        labels = c("Biden Voter", "Trump Voter")),
    prepost    = factor(prepost, levels = c(0, 1),
                        labels = c("Pre-Election", "Post-Election"))
  ) %>%
  ggplot(aes(x = vote_trump, y = .epred, fill = prepost)) +
  geom_split_violin(alpha = 0.6, trim = FALSE) +
  geom_boxplot(width = 0.15, alpha = 0.4,
               position = position_dodge(0.3), outlier.shape = NA) +
  scale_fill_manual(values = c("Pre-Election" = "#457B9D", "Post-Election" = "#E63946"),
                    name = NULL) +
  labs(
    title = "Predicted Support for Ballot Recount",
    x = NULL,
    y = "Posterior Predicted Mean"
  ) +
  mytheme() +
  theme(legend.position = "bottom")
```

### Multi-item split violin panel

```{r split-violin-multi}
all_draws %>%
  ungroup() %>%
  mutate(
    vote_trump = factor(vote_trump, levels = c(0, 1),
                        labels = c("Biden Voter", "Trump Voter")),
    prepost    = factor(prepost, levels = c(0, 1),
                        labels = c("Pre", "Post")),
    item       = factor(item,
                        levels = items,
                        labels = c("Attend March", "Burn Flag", "Recount",
                                   "Go to Court", "Criticize Election"))
  ) %>%
  ggplot(aes(x = item, y = .epred, fill = prepost)) +
  geom_split_violin(alpha = 0.5, trim = FALSE, scale = "width") +
  facet_wrap(~vote_trump, ncol = 1) +
  scale_fill_manual(values = c("Pre" = "#457B9D", "Post" = "#E63946"), name = NULL) +
  labs(title = "Posterior Distributions by Item and Vote Choice",
       x = NULL, y = "Predicted Mean") +
  mytheme() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 30, hjust = 1)
  )
```

## 3. Forest Plots: Predictive Differences

Forest plots display point estimates and credible intervals for treatment effects. `gg_pred_diff()` produces these from posterior predictive difference summaries.

```{r forest-basic}
all_diffs <- map_dfr(items, function(it) {
  posterior_pred_diff(
    model = models[[paste0(it, "_linear")]],
    xvar  = "prepost",
    mvar  = "vote_trump"
  ) %>%
    mutate(item = it)
})

gg_pred_diff(
  data         = all_diffs,
  x_var        = "item",
  color_var    = "vote_trump",
  color_labels = c("0" = "Biden Voter", "1" = "Trump Voter"),
  x_labels     = c(
    "attend_march"       = "Attend March",
    "burn_flag"          = "Burn Flag",
    "recount"            = "Recount",
    "court"              = "Go to Court",
    "criticize_election" = "Criticize Election"
  ),
  title = "Post-Election Effect on Contestation Support"
)
```

### Comparing model families

```{r forest-families}
diffs_both <- map_dfr(items, function(it) {
  bind_rows(
    posterior_pred_diff(
      models[[paste0(it, "_linear")]],
      xvar = "prepost", mvar = "vote_trump"
    ) %>% mutate(item = it, model = "Linear"),
    posterior_pred_diff(
      models[[paste0(it, "_ord")]],
      xvar = "prepost", mvar = "vote_trump",
      category = "5"
    ) %>% mutate(item = it, model = "Ordinal (Pr = 5)")
  )
})

gg_pred_diff(
  data         = diffs_both,
  x_var        = "item",
  color_var    = "vote_trump",
  facet_var    = "model",
  color_labels = c("0" = "Biden", "1" = "Trump"),
  title        = "Effect Comparison: Linear vs. Ordinal"
)
```

## 4. Marginal Effects Plots

### Categorical marginal effects (ordinal models)

`ggMargins_categories()` plots the marginal effect per response category with dodged points and error bars:

```{r marginal-categories}
me_ord <- posterior_pme(
  model  = models[["recount_ord"]],
  xvar   = "prepost",
  mvar   = "vote_trump",
  xrange = c(0, 1),
  mrange = c(0, 1)
)

ggMargins_categories(
  data               = me_ord,
  mvar               = "vote_trump",
  title              = "Marginal Effect: Recount (Ordinal)",
  ytitle             = "Change in Probability (Post − Pre)",
  ylimits            = c(-0.15, 0.15),
  point_color_values = c("#E63946", "#457B9D")
)
```

### Horizontal marginal effects

`plotMarginalEffects()` provides an alternative layout with horizontal error bars:

```{r marginal-horizontal}
me_formatted <- me_ord %>%
  filter(vote_trump == 0) %>%
  rename(category = .category, mean_effect = mean, lower_ci = lower, upper_ci = upper)

plotMarginalEffects(
  me_data = me_formatted,
  title   = "Marginal Effect: Biden Voters",
  category_labels = c("1" = "Strongly Oppose", "2" = "Oppose",
                       "3" = "Neutral", "4" = "Support", "5" = "Strongly Support"),
  x_limits = c(-0.15, 0.15)
)
```

### Battery forest plot

`plot_marginal_effects()` summarizes marginal effects across all items and families:

```{r battery-forest}
me_battery <- marginal_effects_battery(
  model_list = models,
  items      = items,
  treatment  = "prepost",
  moderator  = "vote_trump"
)

plot_marginal_effects(
  me_data = me_battery,
  title   = "Marginal Effect of Post-Election Period",
  colors  = c("Trump" = "#E63946", "Non-Trump" = "#457B9D")
)
```

## 5. Dot Plots

`plotDots()` shows predicted effects as dots with optional connector lines and error bars.

### Predicted probabilities by category

```{r dots-predicted}
pred_ord <- posterior_means(
  model = models[["recount_ord"]],
  xvar  = "prepost",
  mvar  = "vote_trump"
)

plotDots(
  plot_dat        = pred_ord,
  y_var           = "mean",
  x_var           = ".category",
  color_var       = "prepost",
  facet_var       = "vote_trump",
  color_levels    = c(0, 1),
  color_labels    = c("Pre-Election", "Post-Election"),
  color_values    = c("grey50", "black"),
  facet_levels    = c(0, 1),
  facet_labels    = c("Biden Voters", "Trump Voters"),
  category_labels = c("Strongly Oppose", "Oppose", "Neutral",
                       "Support", "Strongly Support"),
  title           = "Predicted Probabilities: Recount",
  y_label         = "Posterior Mean Probability",
  y_limits        = c(0, 0.6)
)
```

### Marginal effects as dot plot with error bars

```{r dots-marginal}
plotDots(
  plot_dat       = me_ord %>% filter(vote_trump == 1),
  y_var          = "mean",
  x_var          = ".category",
  color_var      = NULL,
  show_errorbar  = TRUE,
  reference_line = 0,
  connector_lines = FALSE,
  category_labels = c("Strongly Oppose", "Oppose", "Neutral",
                       "Support", "Strongly Support"),
  title          = "Marginal Effect: Trump Voters",
  y_label        = "Change in Probability",
  y_limits       = c(-0.2, 0.2)
)
```

## 6. Sunflower Plots

Sunflower plots overlay predicted probabilities on observed data distributions using the `vayr` package's sunflower dodge positioning.

```{r sunflower, eval = FALSE}
# Requires preparing prediction summary and observed count data
# See ?plotSunflower for expected data format

# pred_summary: data frame with category, mean_prob, lower_ci, upper_ci, prepost_label
# observed_counts: data frame with category, n, prepost

plotSunflower(
  pred_summary    = pred_summary,
  observed_counts = observed_counts,
  x_var           = "prepost_label",
  join_vars       = c("prepost", "category"),
  title           = "Predicted Probabilities: Recount",
  y_limits        = c(0, 0.5)
)
```

## 7. Combined Panels

`create_combined_plot()` assembles a 2x2 panel with sunflower plots (left) and marginal effects (right) for two moderator groups. This is the publication-ready layout used in the electoral contestation paper.

```{r combined, eval = FALSE}
create_combined_plot(
  plot_data        = plot_data_list,
  me_data          = me_data,
  main_title       = "Support for Ballot Recount",
  moderator        = "vote_trump",
  moderator_levels = c(0, 1),
  group_labels     = c("Biden Voters", "Trump Voters"),
  category_labels  = c("Strongly Oppose", "Oppose", "Neutral",
                        "Support", "Strongly Support"),
  join_vars        = c("category", "prepost"),
  color_labels     = c("Pre-Election", "Post-Election"),
  color_values     = c("grey60", "black")
)
```

## 8. Custom Compositions with patchwork

For full control, compose individual plots with `patchwork`:

```{r patchwork}
# Ridgeplot for one item
p_ridge <- gg_posterior_ridges(
  draws     = ord_draws %>% filter(item == "recount"),
  ridge_var = "prepost",
  fill_var  = "vote_trump",
  category  = "5",
  fill_labels = c("0" = "Biden", "1" = "Trump"),
  title     = "Pr(Strongly Support Recount)"
)

# Forest plot for same item
p_forest <- gg_pred_diff(
  data       = all_diffs %>% filter(item == "recount"),
  x_var      = "vote_trump",
  color_var  = "vote_trump",
  color_labels = c("0" = "Biden", "1" = "Trump"),
  title      = "Post − Pre Difference"
)

# Marginal effects
p_margins <- ggMargins_categories(
  data  = me_ord,
  mvar  = "vote_trump",
  title = "Marginal Effect by Category",
  point_color_values = c("#E63946", "#457B9D")
)

# Compose
(p_ridge | p_forest) / p_margins +
  plot_annotation(
    title   = "Ballot Recount: Post-Estimation Summary",
    theme   = theme(plot.title = element_text(size = 14, face = "bold"))
  )
```

## 9. The `mytheme()` Publication Theme

All `vrm` plots can be styled with `mytheme()` for consistent publication-quality appearance:

```{r mytheme}
# Apply to any ggplot
ggplot(data.frame(x = rnorm(200), y = rnorm(200)), aes(x, y)) +
  geom_point(alpha = 0.4) +
  labs(title = "Publication Theme", x = "X", y = "Y") +
  mytheme()
```

## Summary

| Plot Type | Function | Best For |
|-----------|----------|----------|
| Ridgeplot | `gg_posterior_ridges()` | Full posterior distributions across items |
| Split Violin | Custom `geom_split_violin` | Comparing two distributions side by side |
| Forest Plot | `gg_pred_diff()` | Point estimates with credible intervals |
| Marginal Effects (categorical) | `ggMargins_categories()` | Per-category effects for ordinal models |
| Marginal Effects (horizontal) | `plotMarginalEffects()` | Single-group marginal effects |
| Battery Forest | `plot_marginal_effects()` | Cross-item, cross-family comparisons |
| Dot Plot | `plotDots()` | Predicted values with connectors/error bars |
| Sunflower Plot | `plotSunflower()` | Predictions overlaid on observed counts |
| Combined Panel | `create_combined_plot()` | Publication 2x2 panels |
